---
title: "gravitas"
author: "Sayani Gupta"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{gravitas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!-- #Example vignette: 
https://github.com/njtierney/naniar/blob/master/vignettes/getting-started-w-naniar.Rmd -->
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,   # code
  include    = TRUE,   # plots
  results    = "hide", # text: "hide", "show"
  eval       = TRUE,   # chunk
  message    = FALSE,
  warning    = FALSE,
  error      = FALSE,
  collapse   = TRUE,
  comment    = "#>",
  fig.height = 4,
  fig.width  = 6,
  fig.align  = "center",
  cache      = FALSE
)
library(gravitas)
```

# Introduction


## Dynamic projections of high dimensional data, tours

*Tours* are a class of dynamic orthogonal (linear) projections that embed $p-$dimensional (Euclidean) space into a $d-$dimensional subspace and animates many such projections as the projection basis (orientation) is rotated. Tours are useful in identify clustering, outliers, and structure held within numeric multivariate data. 


## Notation



# Usage



## Example 1 - Rotate an arbitrary basis (& utility functions)



```{r view-basis}

```



```{r view-manip-space}

```



```{r, eval=F}

```

*plotly animation not rendered in this vignette, but will display in viewer*

## Example 2 - Rotating a glyph map



## Example 3 - Rotation after holes guided tour


# Summary



## Namesake 



## Acknowledgements

This package and vignette were made in *R* with much support from `devtools` and `roxygen2`.

Thanks to Prof. Dianne Cook for providing focus, namesake, and the groundwork underpinning this work. Thanks to Dr. Ursula Laa for use-cases and applications.
