---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# gravitas

<!-- badges: start -->

[![Travis-CI Build Status](https://travis-ci.org/Sayani07/gravitas.svg?branch=master)](https://travis-ci.org/Sayani07/gravitas)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/Sayani07/gravitas?branch=master&svg=true)](https://ci.appveyor.com/project/Sayani07/gravitas)
[![Codecov test coverage](https://codecov.io/gh/Sayani07/gravitas/branch/master/graph/badge.svg)](https://codecov.io/gh/Sayani07/gravitas?branch=master)



<!-- badges: end -->

## Overview

The package **gravitas** provides a tool to examine the probability distribution of univariate time series across bivariate temporal granularities using range of graphics in `ggplot2` through the following:  

 - create multiple-order-up circular or aperiodic temporal granularities.  
 
 - categorize pairs of granularities as either  *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.  
 
 - recommending appropriate probability distribution plots of the time series variable across the bivariate granularities based on the levels of the bivariate granularties and their interaction.  

## Installation

You can install gravitas from CRAN:
  
```{r, eval = FALSE}
install.packages("gravitas")  
```

You can install the development version from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("Sayani07/gravitas")

```
## Quick look

gravitas comes with an interactive webpage, which lets you go through the different functionalities of this package. To try it, simply use gravitas::run_app().


## Features

  -  Search for a set of all possible temporal granularities with `search_gran`

  -  Build any temporal granularity with `create_gran` 

  - Refine your search of possible temporal granularities by altering arguments in `search_gran()`

 
  - Check if two temporal granularities are harmonies with `is_harmony()`
 
  - Get all possible harmonies with `harmony()`
 
  - Get recommendations on choosing more appropriate distribution plots and advice on interaction between granularities, number of observations available for drawing probability distributions for chosen granularities with `gran_advice`
  
  - Explore probability distribution across bivariate temporal granularities with `prob_plot()`
    
## Example

The probability distribution of the energy consumption for one household from 
[customer trials](https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/details?q=smart-meter) across weekend/weekday and half-hours of the day can be examined through a quantile plot or letter value plot.

```{r, message=FALSE}
library(gravitas)
library(dplyr)
library(ggplot2)
library(tsibble)

smart_meter <- smart_meter10 %>% as_tsibble() %>% ungroup()

smart_meter %>% 
  prob_plot(gran1 = "wknd_wday",
            gran2 = "hour_day",
            response = "general_supply_kwh",
            plot_type = "quantile",
                       quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  scale_y_sqrt()
```    

```{r}
library(gravitas)
smart_meter %>% 
  prob_plot("wknd_wday",
            "hour_day",
                              response = "general_supply_kwh",
                              plot_type = "lv",
                       outlier.colour = "red") + scale_y_sqrt()
```    
  
## More information

View the vignette to get started! 


```{r}
library("gravitas")
vignette("gravitas_vignette")
```


This package takes tsibble as the data input. Tsibble provides a data class of tbl_ts to represent tidy temporal data. It consists of a time index, key and other measured variables in a data-centric format, which makes it easier to work with temporal data. To learn more about it, please visit [https://tsibble.tidyverts.org/](https://tsibble.tidyverts.org/)

## Reporting and issues
Please submit all bug reports, errors, and feature requests to https://github.com/Sayani07/gravitas/issues





