---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(gravitas)
library(ggplot2)

```
# gravitas

<!-- badges: start -->

[![Travis-CI Build Status](https://travis-ci.org/Sayani07/gravitas.svg?branch=master)](https://travis-ci.org/Sayani07/gravitas)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/Sayani07/gravitas?branch=master&svg=true)](https://ci.appveyor.com/project/Sayani07/gravitas)
[![Codecov test coverage](https://codecov.io/gh/Sayani07/gravitas/branch/master/graph/badge.svg)](https://codecov.io/gh/Sayani07/gravitas?branch=master)



<!-- badges: end -->

## Overview

The package **gravitas** provides a tool to examine the probability distribution of univariate time series across bivariate temporal granularities using range of graphics in `ggplot2` through the following:  

 - create multiple-order-up circular or aperiodic temporal granularities.  
 
 - categorize pairs of granularities as either  *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.  
 
 - recommending appropriate probability distribution plots of the time series variable across the bivariate granularities based on the levels of the bivariate granularties and their interaction.  

The hierarchical structure of time creates a natural nested ordering. For example, hours are nested within days, days within weeks, weeks within months, and so on. Package `gravitas` is not restricted to temporal data. It can be used in non-temporal applications for which a hierarchical structure can be constructed similar to time.


## Installation

You can install gravitas from CRAN:
  
```{r, eval = FALSE}
install.packages("gravitas")  
```

You can install the development version from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("Sayani07/gravitas")

```
## Quick look

gravitas comes with an interactive webpage, which lets you go through the different functionalities of this package. To try it, simply use gravitas::run_app().


## Features

  -  Search for a set of all possible temporal granularities with `search_gran()`
  
  - Build any temporal granularity with `create_gran` 
  
  - Check if two temporal granularities are harmonies with `is_harmony()`
 
  - Get all possible harmonies with `harmony()`
 
  - Get recommendations on choosing more appropriate distribution plots and advice on interaction between granularities, number of observations available for drawing probability distributions for chosen granularities with `gran_advice()`
  
  - Validate if the created granularity matches your already existing column with `validate_gran()`
  
  - Explore probability distribution across bivariate temporal granularities with `prob_plot()`
    
## Example: temporal case

The probability distribution of the energy consumption for ten households from 
[customer trials](https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/details?q=smart-meter) can be explored as follows:

###  Search for granularities

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(lvplot)

 smart_meter10 %>%
   search_gran(highest_unit = "week")
```

### Screen harmonies from the search list

```{r, message=FALSE}
 smart_meter10 %>%
   harmony(
     ugran = "day",
     filter_in = "wknd_wday"
   )
``` 

### Visualize probability distribution of the harmony pair (wknd_wday, hour_day)
 
 
```{r probplot} 
smart_meter10 %>%
 filter(customer_id == 10017936) %>%
   prob_plot(
     gran1 = "wknd_wday",
     gran2 = "hour_day",
     response = "general_supply_kwh",
     plot_type = "quantile",
     quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
     symmetric = FALSE
   ) +
   scale_y_sqrt()
```    

## Example: non-temporal case

`cricket` data set in the package can be explored by explicitly defining a hierarchy table as follows:

#### Visualize granularities for non-temporal data


```{r cricket}
library(tsibble)
cricket_tsibble <- cricket %>%
 mutate(data_index = row_number()) %>%
 as_tsibble(index = data_index)

 hierarchy_model <- tibble::tibble(
   units = c("index", "ball", "over", "inning", "match"),
   convert_fct = c(1, 6, 20, 2, 1)
 )
 cricket_tsibble %>%
   create_gran(
     "ball_over",
     hierarchy_model
   )

   cricket_tsibble %>%
   filter(batting_team %in% c("Mumbai Indians",
                              "Chennai Super Kings"))%>%
   prob_plot("inning", "over",
   hierarchy_model,
   response = "runs_per_over",
   plot_type = "lv")
```

#### Validate if given column in the data set equals computed granularity

```{r validate}
validate_gran(cricket_tsibble,
   gran = "over_inning",
   hierarchy_tbl = hierarchy_model,
   validate_col = "over" )
```


## Acknowledgements

Thanks to PhD supervisors [Prof. Rob J Hyndman](https://robjhyndman.com/), [Prof. Dianne Cook](http://dicook.org/) and [Google Summer of Code 2019](https://summerofcode.withgoogle.com/) mentor [Prof. Antony Unwin](http://rosuda.org/~unwin/) for their support and always leading by example. The fine balance of encouraging me to work on my ideas and stepping in to help when I need has made the devlopment of this package a great learning experience for me.

Moreover, I want to thank my cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for always lending an ear and sharing their wisdom and experience of developing R packages whenever needed. 

## More information

View the [vignette](https://sayani07.github.io/gravitas/) to get started! 

This package takes tsibble as the data input. Tsibble provides a data class of tbl_ts to represent tidy temporal data. It consists of a time index, key and other measured variables in a data-centric format, which makes it easier to work with temporal data. To learn more about it, please visit [https://tsibble.tidyverts.org/](https://tsibble.tidyverts.org/)

## Reporting and issues
Please submit all bug reports, errors, and feature requests to https://github.com/Sayani07/gravitas/issues





